---
title: "R Notebook"
output: html_notebook
---

Reading in data: large data set of 17.3 million observations. UserID are users which gave a rating on a 10 point scale to other users (profileID).

```{r}
df <- read.delim("data/libimseti/ratings.dat", sep = ",", header = F)
colnames(df) <- c("userID", "profileID", "rating")
head(df)
```

Now the data need to be converted to rating matrix. These files contain 17,359,346 anonymous ratings of 168,791 profiles
made by 135,359 LibimSeTi users as dumped on April 4, 2006. Not all profiles gave ratings to others. Rating matrix is created with the 'Matrix' package.
```{r}
library(Matrix)
library(dplyr)
ratingMatrix <- sparseMatrix(i = df$userID,
                         j = df$profileID,
                         x = df$rating)

table_rat <- table(as.vector(ratingMatrix@x))

table <- rbind(table_rat, prop.table(table_rat))
rownames(table) <- c("# Ratings", "% of Ratings")
tab_df <- as.data.frame(table)
bind_rows(round(tab_df[1,], 0), round(tab_df[2,], 2))

```

